SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-12 docs/questioning.tex
fixes up to IV

  Some of the most dangerous web application attacks, like Cross-Site
  Scripting and {\sc sql} injection, exploit the oversight of
  programmers that develop applications which may accept and process
  data of uncertain origin without appropriate validation and / or
  filtering. Such attacks, relying on source code injection
  vulnerabilities, have been constantly topping the lists of the
  various security bulletin providers despite the numerous
  countermeasures that have been proposed over the past 15 years.

  In this paper we systematize the existing knowledge regarding
  various defense mechanisms against source code injection by setting
  up a model that highlights the key weaknesses behind such attacks and
  helps us observe defenses from a common perspective.

  We then categorize and analyze the various defense mechanisms that
  have been developed. Our analysis is based on dimensions such as
  accuracy, performance, deployment, security and availability.
  Especially notable is accuracy, as our findings show that there are
  many defense mechanisms that are tested in a poor manner.
  In addition, we observe that there are mechanisms
  that can be bypassed by attackers that know how the mechanisms
  work.

  By comparing the existing defenses and identifying their common
  characteristics, we help: a) researchers and practitioners to choose
  or modify a mechanism according to their needs, and b) designers of
  new approaches in developing more efficient mechanisms.

@@

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different
source code-driven injection attacks dominate the top five
positions. Malicious users find new ways~\cite{HNSHS12,DKH14}
to bypass defense mechanisms by using a variety of
techniques despite the numerous countermeasures
that are being introduced. Note that the number of systems developed
to counter {\sc sql} injection attacks until 2006 was more than
twenty~\cite{HVO06}. Since then the number has doubled. Researchers
have also indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

In this paper we initially show how the above attacks can be seen from
a {\it common perspective}. We achieve this by providing an
exploitation model that highlights the following fact: the majority of
the steps needed to perform application attacks based on
source code injection are common.\footnote{Note that we leave out
  attacks that are based on binary code injection (i.e. buffer
  overflows --- see Section~\ref{sec:attacks}).} This is validated by
the fact that there are protection mechanisms that can detect more
than one of these attacks. Then, we {\it categorize} a selection of
these protection mechanisms. The selection comprises protection
mechanisms that detect such attacks and are presented in publications
cited more than twenty times, and some of the latest mechanisms that
have been presented in top security conferences that have been cited
less than twenty times. We included the second group so that we would
not penalize recent research. Finally, we {\it analyze} each mechanism based
on the following dimensions:
\begin{itemize}
\item {\it Accuracy} Protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Computational Performance} We examine
  the overhead of the mechanisms
  that affect the experience of a web user who actually
  uses the protected application.
\item {\it Ease of use} We examine if the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security} We examine how resilient is the system to
  attackers that know how it works and wants to circumvent it.
\item {\it Point of Detection} Based on our exploitation model
  we examine the location where a mechanism detects an attack.
\item {\it Availability} We check if the mechanism and its
  testbed are available on-line.
available 
\end{itemize}

@@

There is already a survey on mitigating software vulnerabilities
in general~\cite{SZ12}. The scope of that
research is pretty wide and leaves out the majority of the
approaches and mechanisms that we report here.
This is mainly because this work includes countermeasures
that detect such vulnerabilities via program analysis
(which is done during development or testing), while
we include protection mechanisms that counter
such attacks when they take place. Static or
dynamic analyzers that examine applications to
identify vulnerabilities that may lead to the attacks
that we described earlier are out of the scope of this paper.
Also, countermeasures that prevent {\sc sql}
injection attacks~\cite{HVO06}, have already been surveyed.
However, this survey is quite old and since then
the number of countermeasures that detect {\sc sql}
injection attacks alone, has been drastically increased
as we mentioned earlier.

@@

Bratus et al.~\cite{BLSPS11} portray the issue in a generic fashion:
{\it ``unexpected (and unexpectedly powerful) computational models
  inside targeted systems, which turn a part of the target into a
  so-called `weird machine' programmable by the attacker via crafted
  inputs (a.k.a. `exploits')''}. In particular, {\it ``every
  application that copies untrusted input verbatim into an output
  program is vulnerable to code injection''}. Ray and
Ligatti~\cite{RL12b} actually proved the above claim based on formal
language theory.

@@

Application attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s like {\sc sql} and
{\sc xml} play an significant role in the development of either web or
mobile applications. For example, many applications have interfaces
where a user enters input to interact with the application's data,
thereby interacting with the underlying Relational Database Management
System ({\sc rdbms}). This input can become part of an {\sc sql} query
and executed on the target {\sc rdbms}. A code injection attack that
exploits the vulnerabilities of these interfaces by taking advantage
of input validation issues like incorrectly passed parameters or
incorrect type handling, is called an {\sc sql} injection
attack~\cite{CERT02,MS09,HVO06,SW06}. By using similar techniques
malicious users can perform other exploits based on {\sc dsl}s,
like {\sc xp}ath~\cite{SW06,CDL07,MKS09}, {\sc xml}~\cite{MSM13}
and No{\sc sql}~\cite{SMS13}. With this kind of attacks,
a malicious user can view sensitive information,
destroy or modify protected data, or even crash the entire
application. {\sc html} is another {\sc dsl} that can be used
for malicious purposes when an application does not properly
handle user supplied data. Based on this vulnerability
an attacker can supply valid {\sc html},
typically via a parameter value, and inject their own
content into the page. {\sc html} injection is mainly associated
with {\sc xss} defects~\cite{HNSHS12}.

A recent class of {\sc cia}s involve dynamic languages like
JavaScript and {\sc php}~\cite{SFVM09,EWKK09,SMS13}. JavaScript
injection attacks, in particular, make up a large subset of dynamic
language-driven attacks. Such attacks are manifested when a web
application accepts and redisplays data of uncertain origin without
appropriate validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}. JavaScript injection
attacks are considered as a critical issue in web application security
mainly because they are associated with major vulnerabilities such as:
{\sc xss} attacks~\cite{SG07}, {\sc csrf} attacks~\cite{BJM08,LZRL09} and
Cross-Channel Scripting ({\sc xcs}) attacks~\cite{W10,BBB09}.
In many cases {\sc xss} attacks involve the injection of
both {\sc html} and JavaScript code.

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities.
Figure~\ref{fig:attacks} presents the steps of different application
attacks. For every arrow there are
corresponding transition numbers for every attack.
Every number indicates the number of the
steps taken until we reach this transition.
For the most part, the steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Most importantly,
Figure~\ref{fig:attacks} illustrates in which transitions
the defenses detect corresponding attacks.

A malicious user can initiate the attack through two different routes.
The attacker may use the browser of an innocent user as an attack
vehicle, through which the code will be injected in the application.
For example, the attacker could craft a malicious script into a {\sc url}
and then trick an end user to click on it via social engineering (i.e., phishing ---
{\it transitions}: {\sc p-xss} 1.1, {\sc n-xss} 1.1, {\sc csrf} 1).
Alternatively, the attacker may be able to inject
directly the malicious code in the application. For example, consider
a web application that accepts and redisplays user input
without appropriate validation. An attacker could
upload data containing a specially crafted script (for example, in a
blog comment) to steal the cookies of the visiting users or embed
malicious {\sc sql} code to retrieve all the password entries from a
database ({\it transitions}: {\sc dsl} 1,
{\sc p-xss} 1, {\sc n-xss} 1).
Note that an {\sc xss} attack can start from both routes.

@@

{\sc xss} attacks fall ino two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request ({\it transitions}: {\sc p-xss} 7, {\sc n-xss} 2).
Thus, in a non-persistent {\sc xss} attack the injected code
is never saved on the server and immediately becomes part of the
content that is sent back to a web user.

@@

To ensure replicability of research results the source code
implementing a detection mechanism should be available to researchers.
Ideally the code should be available under an open source license, so
that it is easy to modify and improve an approach; even when this is
not possible, for whatever reason, it is important to make sure that
all computer code and test data is available noting any restrictions
on accessibility. Availability of computer code has been recognised as
an important issue outside the computer science field: the journal
Nature, in a recent editorial, adopted a publication policy stressing
access to code and data~\cite{nature2014}. At a time and date that the
merits of open access to code are discussed in non computer science
journals~\cite{easterbrook2014} we should expect computer scientists
to lead the way.

@@

All detection mechanisms extract a cost for their use, as they
introduce some amount of extra computation on an existing application.
The extra computation depends on the exact defense mechanism. For
example, it may be some form of run-time checking, or some form of
obfuscation. Also depending on the mechanism, the cost may be incurred
on different places: it may take place on a server, affecting service
performance, or it may take place on the client, or both. The
usefulness of a detection mechanism depends therefore on the
computational cost it requires and on where it extracts it, as
different overheads may be acceptable server-side than client-side.
For example, reporting absolute measurements gives little information
on the real overhead, unless separate measurements are given for the
system under study with and without the proposed mechanism. Percentage
measurements are normally better. Outliers may be important. It is
always necessary to describe the experimental setup.

@@

The ease of use, in turn, depends on several factors. The detection
mechanism may be deployed server-side or client-side, or both.
Deployments on just the server or the client are easier to handle than
deployments on both of them. The mechanism may be an add-on, or
plugin, on existing software, client or server, or it may be tightly
integrated with existing software, requiring rebuilding from source
code.

@@

Defenders and attackers are often caught in a cat-and-mouse game,
where countermeasures are bypassed by canny attacks, which are caught
by more sophisticated countermeasures, yet again bypassed by cannier
attacks, and so on. We use Security to refer to the ability of a
detection mechanism to resist being circumvented.

>>>>

  Some of the most dangerous web application attacks, like Cross-Site
  Scripting and {\sc sql} injection, exploit the oversight of
  programmers that develop applications that may accept and process
  data of uncertain origin without appropriate validation and / or
  filtering. Such attacks, relying on source code injection
  vulnerabilities, have been constantly topping the lists of the
  various security bulletin providers despite the numerous
  countermeasures that have been proposed over the past 15 years.

  In this paper we systematize the existing knowledge regarding
  various defense mechanisms against source code injection by setting
  up a model that highlights the key weaknesses behind such attacks
  and provides a common perspective for the different defenses.

  We then categorize and analyze the various defense mechanisms that
  have been developed. Our analysis is based on accuracy, performance,
  deployment, security and availability. Especially notable is
  accuracy, as our findings show that there are many defense
  mechanisms that are tested in a poor manner. In addition, we observe
  that there are mechanisms that can be bypassed by attackers that
  know how the mechanisms work.

  By comparing the existing defenses and identifying their common
  characteristics, we help researchers and practitioners to choose or
  modify a mechanism according to their needs as well as designers of
  new approaches in developing more efficient mechanisms.

@@

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different source
code-driven injection attacks dominate the top five positions.
Malicious users find new ways~\cite{HNSHS12,DKH14} to bypass defense
mechanisms by using a variety of techniques despite the numerous
countermeasures that are being introduced. The number of systems
developed to counter {\sc sql} injection attacks until 2006 was more
than twenty~\cite{HVO06}. Since then the number has doubled.
Researchers have indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

In this paper we show how the above attacks can be seen from a {\it
  common perspective}. We achieve this by providing an exploitation
model that highlights that the majority of the steps needed to perform
application attacks based on source code injection are
common.\footnote{Note that we leave out attacks that are based on
  binary code injection (i.e. buffer overflows --- see
  Section~\ref{sec:attacks}).} This is validated by the fact that
there are protection mechanisms that can detect more than one of these
attacks. Then, we {\it categorize} a selection of these protection
mechanisms. In the selection we included protection mechanisms that
detect such attacks and are presented in publications cited more than
twenty times. We also added some recent mechanisms that have been
presented in top security conferences, even though they have been
cited less than twenty times. We included the second group so that we
would not penalize recent research. Finally, we {\it analyze} each
mechanism based on the following dimensions:
\begin{itemize}
\item {\it Accuracy} Protection mechanisms are as good
  as their detection capability; this requires low false positive and
  false negative rates.
\item {\it Availability} We check if the mechanism and its
  testbed are available on-line.
\item {\it Computational Performance} We examine the computational
  overhead of the mechanisms at their points of deployment.
\item {\it Ease of use} We examine if the protection
  mechanism is practical in terms of deployment
  and can be easily adopted by security experts.
\item {\it Security} We examine how resilient is the system to
  attackers that know how it works and wants to circumvent it.
\item {\it Point of Detection} Based on our exploitation model
  we examine the location where a mechanism detects an attack.
\end{itemize}

@@

There is already a survey on mitigating software vulnerabilities
in general~\cite{SZ12}. The scope of that
research is pretty wide and leaves out the majority of the
approaches and mechanisms that we report here.
This is mainly because that work includes countermeasures
that detect such vulnerabilities via program analysis
(which is done during development or testing), while
we include protection mechanisms that counter
such attacks when they take place. Static or
dynamic analyzers that examine applications to
identify vulnerabilities that may lead to the attacks
that we described earlier are out of the scope of this paper.
Also, countermeasures that prevent {\sc sql}
injection attacks~\cite{HVO06}, have already been surveyed.
However, this survey is quite old and since then
the number of countermeasures that detect {\sc sql}
injection attacks alone, has been drastically increased
as we mentioned earlier.

@@

Bratus et al.~\cite{BLSPS11} portray the issue in a generic fashion:
{\it ``unexpected (and unexpectedly powerful) computational models
  inside targeted systems, which turn a part of the target into a
  so-called `weird machine' programmable by the attacker via crafted
  inputs (a.k.a. `exploits')''}. In particular, {\it ``every
  application that copies untrusted input verbatim into an output
  program is vulnerable to code injection''}. Ray and
Ligatti~\cite{RL12b} have proved this claim based on formal language
theory.

@@

Application attacks that involve {\sc dsl}s constitute an important
subset of the code injection problem, as {\sc dsl}s like {\sc sql} and
{\sc xml} play an significant role in the development of both web and
mobile applications. For example, many applications have interfaces
where a user enters input to interact with the application's data,
thereby interacting with the underlying Relational Database Management
System ({\sc rdbms}). This input can become part of an {\sc sql} query
and get executed on the target {\sc rdbms}. A code injection attack that
exploits the vulnerabilities of these interfaces by taking advantage
of input validation issues like incorrectly passed parameters or
incorrect type handling is called an {\sc sql} injection
attack~\cite{CERT02,MS09,HVO06,SW06}. By using similar techniques
malicious users can perform other exploits based on {\sc dsl}s
like {\sc xp}ath~\cite{SW06,CDL07,MKS09}, {\sc xml}~\cite{MSM13}
and No{\sc sql}~\cite{SMS13}. With this kind of attacks,
a malicious user can view sensitive information,
destroy or modify protected data, or even crash the entire
application. {\sc html} is another {\sc dsl} that can be used
for malicious purposes when an application does not properly
handle user supplied data. Based on this vulnerability
an attacker can supply valid {\sc html},
typically via a parameter value, and inject their own
content into the page. {\sc html} injection is mainly associated
with {\sc xss} defects~\cite{HNSHS12}.

A recent class of code injection attacks involve dynamic languages
like JavaScript and {\sc php}~\cite{SFVM09,EWKK09,SMS13}. JavaScript
injection attacks, in particular, make up a large subset of dynamic
language-driven attacks. Such attacks are manifested when a web
application accepts and redisplays data of uncertain origin without
appropriate validation and filtering. Based on this flaw, an attacker
can manage to inject a script in the JavaScript engine of a browser
and alter its execution flow~\cite{ELX07}. JavaScript injection
attacks are considered a critical issue in web application security
mainly because they are associated with major vulnerabilities such as
{\sc xss} attacks~\cite{SG07}, {\sc csrf} attacks~\cite{BJM08,LZRL09}
and Cross-Channel Scripting ({\sc xcs}) attacks~\cite{W10,BBB09}. In
many cases {\sc xss} attacks involve the injection of both {\sc html}
and JavaScript code.

@@

In this section we provide a step-by-step exploitation model to
understand the process of carrying out application attacks based on
source code injection vulnerabilities. Figure~\ref{fig:attacks}
presents the steps of different application attacks. Links between
steps are labelled with the paths constituting different attacks. The
sequence of steps for each attack is shown by the numbers next to the
attack names in the labels. Most steps are common in all attacks. This
is because they are based on similar attack vectors as we mentioned in
the previous section. Large crosses on transitions in
Figure~\ref{fig:attacks} indicate the points where defenses detect
attacks.

A malicious user can initiate the attack through two different routes.
The attacker may use the browser of an innocent user as an attack
vehicle, through which the code will be injected in the application.
For example, the attacker could craft a malicious script into a {\sc
  url} and then trick an end user to click on it via social
engineering (i.e., phishing, through transition {\sc p-xss} 1.1, {\sc
  n-xss} 1.1, {\sc csrf} 1). Alternatively, the attacker may be able
to inject directly the malicious code in the application (through
transition {\sc dsl} 1, {\sc p-xss} 1, {\sc n-xss} 1). This would happen
in a web application that accepts and redisplays user input
without appropriate validation. An attacker could upload data
containing a specially crafted script (for example, in a blog comment)
to steal the cookies of the visiting users or embed malicious {\sc
  sql} code to retrieve all the password entries from a database. Note that
an {\sc xss} attack can start from both routes.

@@

{\sc xss} attacks fall into two categories, {\it non-persistent} and
{\it persistent}. Non-persistent {\sc xss} attacks take place when the
data provided by a user is used immediately by server-side scripts to
parse and display a page of results, without properly sanitizing the
request (transition {\sc p-xss} 7, {\sc n-xss} 2).
Thus, in a non-persistent {\sc xss} attack the injected code
is never saved on the server and immediately becomes part of the
content that is sent back to a web user.

@@

To ensure replicability of research results the source code
implementing a detection mechanism should be available to researchers.
Ideally the code should be available under an open source license, so
that it is easy to modify and improve an approach; even when this is
not possible, for whatever reason, it is important to make sure that
all computer code and test data is available noting any restrictions
on accessibility. Availability of computer code has been recognised as
an important issue outside the computer science field: the journal
\emph{Nature}, in a recent editorial, adopted a publication policy stressing
access to code and data~\cite{nature2014}. At a time and date that the
merits of open access to code are discussed in non computer science
journals~\cite{easterbrook2014}, we should expect computer scientists
to lead the way.

@@

All detection mechanisms extract a cost for their use, as they
introduce some amount of extra computation on an existing application.
The extra computation depends on the exact defense mechanism. For
example, it may be some form of run-time checking, or some form of
obfuscation. Also depending on the mechanism, the cost may be incurred
on different places: it may take place on a server, affecting service
performance, or it may take place on the client, or both. The
usefulness of a detection mechanism depends therefore on the
computational cost it requires and on where it extracts it, as
different overheads may be acceptable server-side than client-side.

Other factors matter as well. Reporting absolute measurements gives
little information on the real overhead, unless separate measurements
are given for the system under study with and without the proposed
mechanism. Percentage measurements are normally better. Outliers may
be important. It is always necessary to describe the experimental
setup.

@@

Ease of use depends on the deployment protocol required for the
mechanism. The detection mechanism may be deployed server-side or
client-side, or both. Deployments on just the server or the client are
easier to handle than deployments on both of them. The mechanism may
be an add-on, or plugin, on existing software, client or server, or it
may be tightly integrated with existing software, requiring rebuilding
from source code.

@@

Defenders and attackers are often caught in a cat-and-mouse game,
where countermeasures are bypassed by canny attacks, which are caught
by more sophisticated countermeasures, yet again bypassed by cannier
attacks, and so on. We use security to refer to the ability of a
detection mechanism to resist being circumvented.
